Computation and comparison of the stable Northeastern US marine boundary layer

### 2020-12-03 Paper/results.tex
Finished making Alan's changes

The suitability of these meshes and domain sizes were determined based
on the resulting wind spectra and turbulent correlation metrics from
the simulated ABL. The wind spectra $S_i(f)$ is defined as a function
of the frequency $f$
\begin{equation}
  \int_0^\infty S_i(f) \textrm{d}f = \sigma_i^2
\end{equation}
where $\sigma_i$ is the wind speed variance and the index $i=u,v,w$
denotes the longitudinal, lateral, or vertical velocity, respectively.
For the LES simulations in this study, we desire that that the mesh
resolution be sufficiently fine to capture the high frequency
components of the spectra past the spectral peak.  This helps ensure
that the unsteady characteristics in the boundary layer are resolved.

@@

\begin{equation}
  f_{Ny} = \frac{1}{2\times \textrm{dt}}
\end{equation}
Secondly, the maximum resolvable frequency of convecting fine scale
eddies on a mesh with grid size $\Delta$ and average horizontal wind
speed $\bar{U}_{horiz}$ can be estimated as
\begin{equation}
  \label{eq:fmax}
  f_{max} = \frac{0.6\bar{U}_{horiz}}{N\Delta}.
\end{equation}
Here, equation \ref{eq:fmax} assumes that the turbulent eddies convect
with a velocity of $0.6\bar{U}_{horiz}$ and $N=8$ is chosen for the
purposes of this study.  Due to the alignment of the flow with the
mesh directions, the grid size $\Delta = \sqrt{2}\cdot \textrm{dx}$.
In all cases considered here, $f_{max} < f_{Ny}$, so the limiting
constraint for resolving the high frequency components was the mesh
resolution.

The comparison of the wind spectra and maximum resolvable frequency
$f_{max}$ for the different mesh resolutions is shown in figure
\ref{fig:GridStudySpectra}.  Here spectra is shown at the $z$=20m
height and averaged using 1/3 octave-bins to enchance clarity. For the
$S_u$ and $S_w$ spectra, there noticeable shifts in the peak
amplitude, while the peak amplitudes for $S_v$ remained fairly
constant between cases.  In all cases, the peak frequency of $S_i(f)$
also increased as the resolution increased. While the mesh resolution
was sufficient to capture the peak frequency for $S_u$, the fine
resolution with 2.5m mesh cells was required to capture the spectral
peaks in the lateral and vertical directions.

@@

%%%%%%%%%%% Grid resolution Rij correlation figure %%%%%%%%%%%%%%%%%%%%
% Created in Postprocessing/GridStudy/GridStudy_Rij.ipynb
\begin{figure}%[hbt!]
  \centering
  \includegraphics[width=2.5in]{figures/GridStudy_Rij_Longitudinal.png}
  \includegraphics[width=2.5in]{figures/GridStudy_Rij_Lateral.png}
  \caption{\label{fig:GridStudyRij} Calculation of the averaged
    longitudinal and lateral $\langle R_{11}(\boldsymbol{\xi})
    \rangle$ coefficient at $z$=20m for LES of stable 5m/s case with
    different resolutions.}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The results shown in figure \ref{fig:GridStudyRij} show that the
finest resolution is required to resolve both the longitudinal and the
lateral length scale.  In contrast to the neutral and unstable ABL
cases, the turbulent structures decorrelate rapidly and are noticeably
finer in the lateral direction compared to the longitudinal direction.
While 2.5m resolution is adequate to resolve the longitudinal
turbulent structures, the lateral direction may require even finer
resolution to fully resolve.  This may be an important consideration
in situations where turbulence in a stable boundary layer interacts
with turbine features, such as blade motions or downstream wakes.
However, the additional refinement requires careful balancing against
the additional computational expense in the simulation.

@@

Before examining the behavior the offshore ABL across all wind speeds,
we first compare the results for a single case using both LES codes.
In this section, we consider the stable 5m/s case, calculated in the
same manner using both AMR-Wind and Nalu-Wind.  Both LES codes used a
mesh resolution of 2.5m based on the guidance from section
\ref{sec:gridstudy} and the computational resources available.

@@

The comparison of the horizontally averaged wind profiles is shown in
figure \ref{fig:CompareAMRvsNaluWind_WSDir}.  At the forcing height
z=20m, the wind speed and wind shear match are close to an exact match
between the two LES codes.  The horizontal wind speeds continue to
agree well for $z$<100m, beyond which a relatively constant velocity
difference is observed.   For both AMR-Wind and Nalu-Wind, an
approximately linear veer profile was observed until $z\approx$ 250m,
at which point the wind direction remained constant until the
inversion layer height.

@@

%%%%%%%%%%% Compare Nalu/AMR WS/WDir profiles %%%%%%%%%%%%%%%%%%%%%%%
% Postprocessing/ABLStats/AMRWind_NaluWind_stable_05ms_mesh2_5x2_5x2_5_paper.ipynb
\begin{figure} %[hbt!]
  \centering
  \includegraphics[width=2.5in]{figures/Compare_AMRWind_NaluWind/AMRWind_NaluWind_stable_05ms_mesh2p5_2p5_2p5_WS.png}
  \includegraphics[width=2.5in]{figures/Compare_AMRWind_NaluWind/AMRWind_NaluWind_stable_05ms_mesh2p5_2p5_2p5_Wdir.png}\\
  \caption{\label{fig:CompareAMRvsNaluWind_WSDir} Comparison of
    AMR-Wind and Nalu-Wind velocity profiles for the stable 5 m/s
    boundary layer. The dashed horizontal line corresponds to the
    measurement height z=20m. }
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

For the 5 m/s SBL case AMR-Wind had a median time of 0.266 seconds per time step.
This means the entire simulation took about 6 hours to simulate 20k seconds
of physical time. For this case a total of $288\times 288 \times 384$ cells were used
which were solved in parallel using 972 MPI ranks. Simulations were carried out NREL's
Eagle supercomputer and AMR-Wind was compiled using Intel 2018.
Typically on CPU's we see a strong scaling limit at around 15k cells per MPI rank.
But for this case we used 32,768 cells per MPI rank which is about half the strong scaling limit.
The time per degree of freedom for AMR-Wind is
\[\frac{n_r t_{ts} }{n_{dofs}} = 8\times10^{-6} [s]\]
where $n_r$ is the number of MPI ranks, $t_{ts}$ is time per time step, and $n_{dofs}$ is
the total number of degrees of freedom (for AMR-Wind we are using the number of cells).
This value allows us to compare speed relative to Nalu-Wind when using a different
grid and number of MPI ranks, where a smaller value means a faster time to solution.

In comparison the same case for Nalu-Wind takes 3.404363 seconds per time step which
takes a total of 75.65 hours to complete the simulation. This simulation was carried out
on 1224 MPI ranks on a grid that contains $300\times 300 \times 400$ cells. The equivalent
time per degree of freedom for Nalu-Wind is $1.032 \times 10^{-4}$ s.

@@

The qualitative behavior of the stable boundary layers can be seen
through snapshots of the flow taken from each of the simulations.  In
figure \ref{fig:SnapshotsZ20}, instantaneous snapshots of the
horizontal and vertical velocities for each of the three cases at the
z=20m measurement height are shown.  The largest difference can be
seen in the size of the flow structures between the very unstable and
stable cases.  Longer streaks are evident in the 15 m/s case compared
to the 5 m/s and 15 m/s cases, which is consistent with the nearly
50\% larger integral lengthscale $L$ given in table
\ref{tab:CompareAMRallWS}.  The differences in the vertical velocity
are also evident in \ref{fig:SnapshotsZ20}.  At higher wind speeds the
magnitude of the vertical $w'$ fluctuations increase, which plays a
strong role in the growth the boundary layer.

@@

This observation can also be seen in images taken along the flow
direction $\theta=225^\circ$, as shown in figure
\ref{fig:SnapshotsSide}.  At the time t=20,000 seconds, the growth the
of 15 m/s stable boundary layer is considerably larger than the very
stable 5 m/s and 10 m/s cases, and has nearly reached the inversion
layer height.  To reach the same level of boundary layer growth, the
very stable ABL cases would require longer simulation times, possibly
longer than the nighttime conditions measured experimentally.

@@

Horizontally averaged flow profiles shown in figures
\ref{fig:CompareAMRallWS} and \ref{fig:CompareAMRallTTI} provide a
more quantitative assessment of the ABL evolution.  The rapid growth
of the 15 m/s stable case leads to strong shear throughout the
boundary layer up to the inversion height, while the very stable 5 m/s
and 10 m/s cases showed minimal shear at higher elevations.  The
change in wind direction is also indicative of the boundary layer
penetration.  Approximately linear veer profiles are visible up to
heights of $z\approx$250 m and $z\approx$350 m for the 5 m/s and 10
m/s cases, respectively, while the linear veer is present in the 15
m/s stable case up to $z$=650m.

The behavior of the TI and temperature profiles in figure
\ref{fig:CompareAMRallTTI} closely follows the shear and veer profiles
discussed above.  Despite all cases reaching a similar TI levels at
$z$=20m, the turbulence decays at different rates for each case until
reaching the top of the boundary layer.  The effect of the different
surface temperature changes is also visible in the horizontally
averaged temperature profiles.  Both the 10 m/s and 15 m/s impose
similar temperature changes (-1.4 K/hr and -1.5 K/hr, respectively)
and each similar temperatures at the ocean surface.  However, the
cooling effect in the very stable 10 m/s case is limited to $z$<350
m, similar to the veer and turbulence profiles.

@@

Strong differences based on the atmospheric stratification can also be
seen when we compare the stable ABL cases with the neutral and
unstable counterparts from \cite{cheung2020large}.  In the previous
study, both neutral and unstable offshore boundary layer cases were
considered at the same 5 m/s, 10 m/s and 15 m/s wind speeds.  Similar
to the findings in \ref{sec:stableABLStats}, the 5 m/s and 10 m/s
unstable cases were classified as ``very unstable'', while the 15 m/s
was identified as ``stable'' based on the Obukhov length scale.

The impact of the stability on the turbulent structures can be seen
through the correlation statistics and calculated lengthscales in
figure \ref{fig:AllStabilityRij} and
\ref{fig:AllStabilityLengthscale}.  The large disparity between the
size of the turbulent structures in the stable and neutral or unstable
cases is immediately evident from the calculated $\langle
R_{11}(\boldsymbol{\xi})\rangle$ coefficients.  The longitudinal
lengthscales for the unstable ABL cases can also be an order of
magnitude larger than the stable ABL cases.  The large lateral
lengthscales for the very unstable ABL cases, compared to the lateral
lengthscales for the neutral and stable cases, is also worth noting.
For the 5 m/s and 10 m/s very unstable ABL cases, the lateral
dimensions of the turbulent structures are similar to the longitudinal
lengthscale.  This is consistent with the large convective structures
which develops in those cases.  For the unstable 15 m/s case, the
lateral lengthscales was closer to neutral and stable cases.

%%%%%%%%%%% All stability correlation figure %%%%%%%%%%%%%%%%%%%%
% created in Postprocessing/ABLLength/CompareAll_ABL_Lengthscales.ipynb
\begin{figure}[hbt!]
  \centering
  \includegraphics[width=3in]{figures/AllStability_Rij_Longitudinal.png}
  \includegraphics[width=3in]{figures/AllStability_Rij_Lateral.png}
  \caption{ \label{fig:AllStabilityRij} Calculation of the averaged
    longitudinal and lateral $\langle R_{11}(\boldsymbol{\xi})\rangle$
    coefficient at $z$=20m for all ABL stability cases.}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

>>>>

The suitability of these meshes and domain sizes were determined based
on the resulting wind spectra and turbulent correlation metrics from
the simulated ABL. The wind spectra $S_i(f)$ is defined as a function
of the frequency $f$
\begin{equation}
  \int_0^\infty S_i(f) \textrm{d}f = \sigma_i^2
\end{equation}
where $\sigma_i$ is the wind speed variance and the index $i=u,v,w$
denotes the longitudinal, lateral, or vertical velocity, respectively.
For the LES simulations in this study, we desire that the mesh
resolution be sufficiently fine to capture the high frequency
components of the spectra past the spectral peak.  This helps ensure
that the unsteady characteristics in the boundary layer are resolved.

@@

\begin{equation}
  f_{Ny} = \frac{1}{2\times \textrm{dt}}
\end{equation}
Secondly, the maximum resolvable frequency of convecting fine scale
eddies on a mesh with grid size $\Delta$ and average horizontal wind
speed $\bar{U}_{horiz}$ can be estimated as
\begin{equation}
  \label{eq:fmax}
  f_{max} = \frac{0.6\bar{U}_{horiz}}{N\Delta}.
\end{equation}
Here, Eq. \ref{eq:fmax} assumes that the turbulent eddies convect
with a velocity of $0.6\bar{U}_{horiz}$ and $N=8$ is chosen for the
purposes of this study.  Due to the alignment of the flow with the
mesh directions, the grid size is defined as $\Delta = \sqrt{2}\cdot \textrm{dx}$.
In all cases considered here, $f_{max} < f_{Ny}$, so the limiting
constraint for resolving the high frequency components was the mesh
resolution.

The comparison of the wind spectra and maximum resolvable frequency
$f_{max}$ for the different mesh resolutions is shown in figure
\ref{fig:GridStudySpectra}.  Here spectra is shown at the $z$=20m
height and averaged using 1/3 octave-bins to enhance clarity. For the
$S_u$ and $S_w$ spectra, there are noticeable shifts in the peak
amplitude, while the peak amplitudes for $S_v$ remained fairly
constant between cases.  In all cases, the peak frequency of $S_i(f)$
also increased as the resolution increased. While the mesh resolution
was sufficient to capture the peak frequency for $S_u$, the fine
resolution with 2.5m mesh cells was required to capture the spectral
peaks in the lateral and vertical directions.

@@

%%%%%%%%%%% Grid resolution Rij correlation figure %%%%%%%%%%%%%%%%%%%%
% Created in Postprocessing/GridStudy/GridStudy_Rij.ipynb
\begin{figure}%[hbt!]
  \centering
  \includegraphics[width=2.5in]{figures/GridStudy_Rij_Longitudinal.png}
  \includegraphics[width=2.5in]{figures/GridStudy_Rij_Lateral.png}
  \caption{\label{fig:GridStudyRij} Calculation of the averaged
    longitudinal and lateral $\langle R_{11}(\boldsymbol{\xi})
    \rangle$ coefficients at $z$=20m for LES of stable 5m/s case with
    different resolutions.}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The results shown in figure \ref{fig:GridStudyRij} show that the
finest resolution is required to resolve both the longitudinal and the
lateral length scale.  In contrast to the neutral and unstable ABL
cases, the turbulent structures decorrelate rapidly and are noticeably
finer in the lateral direction compared to the longitudinal direction.
While 2.5m resolution is adequate to resolve the longitudinal
turbulent structures, the lateral direction may require even finer
resolution to fully resolve.  This may be an important consideration
in situations where turbulence in a SBL interacts
with turbine features, such as blade motions or downstream wakes.
However, the additional refinement requires careful balancing against
the additional computational expense in the simulation.

@@

Before examining the behavior of the offshore ABL across all wind speeds,
we first compare the results for a single case using both LES codes.
In this section, we consider the stable 5m/s case, calculated in the
same manner using both AMR-Wind and Nalu-Wind.  Both LES codes used a
mesh resolution of 2.5m based on the guidance from section
\ref{sec:gridstudy} and the computational resources available.

@@

The comparison of the horizontally averaged wind profiles is shown in
figure \ref{fig:CompareAMRvsNaluWind_WSDir}.  At the forcing height
z=20m, the wind speed and wind shear are close to an exact match
between the two LES codes.  The horizontal wind speeds continue to
agree well for $z$<100m, beyond which a relatively constant velocity
difference is observed.   For both AMR-Wind and Nalu-Wind, an
approximately linear veer profile was observed until $z\approx$ 250m,
at which point the wind direction remained constant until the
inversion layer height.

@@

%%%%%%%%%%% Compare Nalu/AMR WS/WDir profiles %%%%%%%%%%%%%%%%%%%%%%%
% Postprocessing/ABLStats/AMRWind_NaluWind_stable_05ms_mesh2_5x2_5x2_5_paper.ipynb
\begin{figure} %[hbt!]
  \centering
  \includegraphics[width=2.5in]{figures/Compare_AMRWind_NaluWind/AMRWind_NaluWind_stable_05ms_mesh2p5_2p5_2p5_WS.png}
  \includegraphics[width=2.5in]{figures/Compare_AMRWind_NaluWind/AMRWind_NaluWind_stable_05ms_mesh2p5_2p5_2p5_Wdir.png}\\
  \caption{\label{fig:CompareAMRvsNaluWind_WSDir} Comparison of
    AMR-Wind and Nalu-Wind horizontal velocity and wind direction
    profiles for the stable 5 m/s boundary layer. The dashed
    horizontal line corresponds to the measurement height z=20m. }
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

For the 5 m/s SBL case AMR-Wind had a median time of 0.266 seconds per time step.
This means the entire simulation took about 6 hours to simulate 20k seconds
of physical time. For this case a total of $288\times 288 \times 384$ cells were used
which were solved in parallel using 972 MPI ranks. Simulations were carried out using NREL's
Eagle supercomputer and AMR-Wind was compiled using Intel 2018.
Typically on CPU's we see a strong scaling limit at around 15k cells per MPI rank.
But for this case we used 32,768 cells per MPI rank which is about half the strong scaling limit.
The time per degree of freedom for AMR-Wind is
\[\frac{n_r t_{ts} }{n_{dofs}} = 8\times10^{-6} [s]\]
where $n_r$ is the number of MPI ranks, $t_{ts}$ is time per time step, and $n_{dofs}$ is
the total number of degrees of freedom (for AMR-Wind we are using the number of cells).
This value allows us to compare speed relative to Nalu-Wind when using a different
grid and number of MPI ranks, where a smaller value means a faster time to solution.

In comparison, the same case for Nalu-Wind takes 3.404363 seconds per time step which
takes a total of 75.65 hours to complete the simulation. This simulation was carried out
on 1224 MPI ranks on a grid that contains $300\times 300 \times 400$ cells. The equivalent
time per degree of freedom for Nalu-Wind is $1.032 \times 10^{-4}$ s.

@@

The qualitative behavior of the stable boundary layers can be seen
through snapshots of the flow taken from each of the simulations.  In
figure \ref{fig:SnapshotsZ20}, instantaneous snapshots of the
horizontal and vertical velocities for each of the three cases at the
z=20m measurement height are shown.  The largest difference can be
seen in the size of the flow structures between the very unstable and
stable cases.  Longer streaks are evident in the 15 m/s case compared
to the 5 m/s and 15 m/s cases, which is consistent with the nearly
50\% larger integral lengthscale $L$ given in table
\ref{tab:CompareAMRallWS}.  The differences in the vertical velocity
are also evident in figure \ref{fig:SnapshotsZ20}.  At higher wind speeds the
magnitude of the vertical $w'$ fluctuations increase, which plays a
strong role in the growth of the boundary layer.

@@

This observation can also be seen in images taken along the flow
direction $225^\circ$, as shown in figure
\ref{fig:SnapshotsSide}.  At the time t=20,000 seconds, the growth of the
of 15 m/s stable boundary layer is considerably larger than the very
stable 5 m/s and 10 m/s cases, and has nearly reached the inversion
layer height.  To reach the same level of boundary layer growth, the
very stable ABL cases would require longer simulation times, possibly
longer than the nighttime conditions measured experimentally.

@@

Horizontally averaged flow profiles shown in figures
\ref{fig:CompareAMRallWS} and \ref{fig:CompareAMRallTTI} provide a
more quantitative assessment of the ABL evolution.  The rapid growth
of the 15 m/s stable case leads to strong shear throughout the
boundary layer up to the inversion height, while the very stable 5 m/s
and 10 m/s cases showed minimal shear at higher elevations (figure \ref{CompareAMRallWS}).  The
change in wind direction is also indicative of the boundary layer
penetration.  Approximately linear veer profiles are visible up to
heights of $z\approx$250 m and $z\approx$350 m for the 5 m/s and 10
m/s cases, respectively, while the linear veer is present in the 15
m/s stable case up to $z$=650m.

The behavior of the TI and temperature profiles in figure
\ref{fig:CompareAMRallTTI} closely follows the shear and veer profiles
discussed above.  Despite all cases reaching a similar TI levels at
$z$=20m, the turbulence decays at different rates for each case until
reaching the top of the boundary layer.  The effect of the different
surface temperature changes is also visible in the horizontally
averaged temperature profiles.  Both the 10 m/s and 15 m/s impose
similar temperature changes (-1.4 K/hr and -1.5 K/hr, respectively)
and similar temperatures at the ocean surface.  However, the
cooling effect in the very stable 10 m/s case is limited to $z$<350
m, similar to the veer and turbulence profiles.

@@

Strong differences based on the atmospheric stratification can also be
seen when we compare the stable ABL cases with the neutral and
unstable counterparts from \cite{cheung2020large}.  In the previous
study, both neutral and unstable offshore boundary layer cases were
considered at the same 5 m/s, 10 m/s and 15 m/s wind speeds.  Similar
to the findings in \ref{sec:stableABLStats}, the 5 m/s and 10 m/s
unstable cases were classified as ``very unstable'', while the 15 m/s
was identified as ``unstable'' based on the Obukhov length scale.

The impact of the stability on the turbulent structures can be seen
through the correlation statistics and calculated lengthscales in
figure \ref{fig:AllStabilityRij} and
\ref{fig:AllStabilityLengthscale}.  The large disparity between the
size of the turbulent structures in the stable and neutral or unstable
cases is immediately evident from the calculated $\langle
R_{11}(\boldsymbol{\xi})\rangle$ coefficients.  The longitudinal
lengthscales for the unstable ABL cases can also be an order of
magnitude larger than the stable ABL cases.  The large lateral
lengthscales for the very unstable ABL cases, compared to the smaller lateral
lengthscales for the neutral and stable cases, is also worth noting.
For the 5 m/s and 10 m/s very unstable ABL cases, the lateral
dimensions of the turbulent structures are similar to the longitudinal
lengthscale.  This is consistent with the large convective structures
which develop in those cases.  For the unstable 15 m/s case, the
lateral lengthscales were closer to neutral and stable cases.

%%%%%%%%%%% All stability correlation figure %%%%%%%%%%%%%%%%%%%%
% created in Postprocessing/ABLLength/CompareAll_ABL_Lengthscales.ipynb
\begin{figure}[hbt!]
  \centering
  \includegraphics[width=3in]{figures/AllStability_Rij_Longitudinal.png}
  \includegraphics[width=3in]{figures/AllStability_Rij_Lateral.png}
  \caption{ \label{fig:AllStabilityRij} Calculation of the averaged
    longitudinal and lateral $\langle R_{11}(\boldsymbol{\xi})\rangle$
    coefficient at $z$=20m for all ABL stability cases and windspeeds.}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
